\documentclass[a4paper]{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[italian]{babel}



% Include any extra LaTeX packages required
\usepackage{xurl}
%\usepackage[hyphens]{url}
\usepackage{hyperref}
\hypersetup{breaklinks=true}

%\usepackage{color}
\usepackage[table,xcdraw,rgb,svgnames]{xcolor}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage[T1]{fontenc}
\usepackage[square, numbers, comma, sort&compress]{natbib}  % Use the "Natbib" style for the references in the Bibliography
\usepackage{verbatim}  % Needed for the "comment" environment to make LaTeX comments
%%\usepackage{vector}  % Allows "\bvec{}" and "\buvec{}" for "blackboard" style bold vectors in maths
\definecolor{airforceblue}{rgb}{0.36, 0.54, 0.66}
\definecolor{ForestGreen}{rgb}{0.13, 0.55, 0.13}
\definecolor{orange}{rgb}{1.0, 0.5, 0.0}
\definecolor{amethyst}{rgb}{0.6, 0.4, 0.8}
\definecolor{awesome}{rgb}{1.0, 0.13, 0.32}
\definecolor{chromeyellow}{rgb}{1.0, 0.65, 0.0}
\hypersetup{colorlinks=true, linkcolor=amethyst, citecolor= ForestGreen, urlcolor=blue }  % Colours hyperlinks in blue, but this can be distracting if there are many links.
\usepackage{quoting}
\quotingsetup{font=small}
\usepackage[output-decimal-marker={,}]{siunitx}
\usepackage{amsfonts} % for the \checkmark command 
\usepackage{listings}
%\usepackage{chemformula}
%\usepackage[newfloat]{minted} % Requires pygments to be installed (Python-based)
\usepackage{sidecap}
\usepackage{booktabs}
\usepackage{multirow} 
\usepackage{caption} 
\usepackage{wrapfig} 
\usepackage{lipsum}  
\usepackage{subfig}  
\usepackage{tabularx}  
\usepackage{longtable}  
\usepackage{microtype}
\usepackage{amsmath}


\newcommand{\vect}[1]{\mathbf{#1}}

\title{Source-aware soundfield estimation based on Physics-constrained Kernel interpolation}
\author{Mattia Marella}
\begin{document}
\maketitle
\section{Introduction}
Sound-field estimation aims at reconstructing the complex acoustic pressure 
$p(\vect{x},\omega)$ in a region of space from a limited number of microphone measurements. This problem arises in spatial audio, sound-zone control, active noise control, and acoustic holography. When only few microphones are available, classical interpolation methods often suffer from spatial aliasing, poor generalization, and strong dependence on microphone placement.

Physics-informed kernel methods provide a powerful alternative: instead of learning an arbitrary regression function, the estimator is constrained to satisfy (at least approximately) the Helmholtz equation, which governs monochromatic acoustic wave propagation:
\[
(\nabla^2 + k^2)p(\mathbf{x}) = 0, \quad k = \frac{\omega}{c}.
\]
Here, $k$ is the wavenumber, $\omega$ is the angular frequency, and $c$ is the speed of sound. 
By incorporating this physical knowledge into the kernel design, these methods can achieve better accuracy and robustness with fewer measurements.

\bigskip
Koyama et al.\ introduced \emph{Physics-Constrained Neural Kernels} (PCNK), where the spatial kernel consists of a superposition of plane waves whose weights are produced by a neural network, but the kernel itself still satisfies the Helmholtz PDE\@. This hybrid approach combines (i) the universal approximation capability of neural networks, and (ii) strong physics priors that regularize learning and reduce overfitting.

The work considered here extends this framework by making the kernel source-aware, meaning that the position of the sound source is explicitly incorporated into the kernel structure. The idea is that acoustic fields are not arbitrary solutions of the Helmholtz equation: they originate from actual sources, and their spatial characteristics (directivity, decay, wavefront orientation) depend on the relative geometry between source and receivers.
So we want the kernel functions to take advantage of this additional information, in order to obtain superior generalization power.


\input{sections/preliminaries.tex}
\input{sections/related_work.tex}
\input{sections/method.tex}
\input{sections/experiment.tex}
\input{sections/training.tex}

\end{document}
